Parsing Postfix log files

### 2008-05-01 doc/logparser-other-parsers-review.tex
Added cluster_groups to the rules table description and example rule. Improved Splunk review. Improved summary table in review of other parsers conclusion. Minor improvements in parser reviews. Improved review of other parsers conclusion.

Splunk aims to index all an organisation's logs, providing a centralised
view capable of searching and correlating diverse log sources.  The web
interface supports complicated searches, providing statistics and graphs in
real time, a facility not provided by \parsername{} (report generation has
been deferred to a subsequent program).  Saved searches\footnote{The author
was unable to save searches, though that may have been due to limitations
in the free version.} can be run periodically and the results emailed to a
recipient or sent to a shell script, which presumably can publish the
results as required (though possibly without the graphs and detailed
statistics); \parsername{} provides the database and leaves it to the user
to utilise it in any way, whenever they want, with no restrictions on usage
of the data.  The interface is optimised for interactive rather than
automated queries and it does not appear to be possible to write
independent tools to utilise the Splunk database; there are additional
reports available on \url{http://www.splunkbase.com/}, but the author was
unable to find any --- every category was empty, even those which the
interface claimed had reports available.  Many types of reports are bundled
with the software, though most are variations of a bar or pie chart, with
the exception of bubble and heatmap graphs.  It is very easy to drill down
through the graphs to extract a portion of the data (e.g.\ select the hour
with the largest number of events, then select a particular host, and
finally a specific address), though it is not possible to search on partial
words.

@@

@@

\subsubsection{Log analyser --- throughput monitor}

@@

\url{http://home.uninet.ee/~ragnar/throughput_monitor/} \newline (Last
checked 2007/08/13.)

@@

YAALA uses a plugin based system to analyse log files and produce \HTML{}
output reports, with all the parsing and report generation handled by
modules.  Using YAALA as a base would be only slightly less work than
starting from scratch, as both input and output modules would need to be
written specially; it may even be more work to implement the parser within
the constraints of YAALA\@.  YAALA supports storing previously gathered
data using Perl's Storable module~\cite{perl-storable}, so with enough
knowledge of the data structure YAALA stores it should be possible to
extract data with another Perl program also using the Storable module;
\parsername{} uses a well documented database which is accessible from the
majority of programming languages.

@@

\subsubsection{Logparser/Lire}

@@

\begin{table}[ht]
\begin{tabular}{l|l|l|l|l}
    Parser              & Parsed     & Data        & Custom        & Compressed \\
                        & test logs? & store?      & reports?      & logs       \\
    \hline                                                                      \\
    Pflogsumm           & Yes        & No          & No            & No         \\
    Sawmill             & Yes        & Yes         & Searches      & Partially  \\
    Splunk              & Yes        & Yes         & Searches      & Yes        \\
                        &            &             & \& reports    &            \\
    Isoqlog             & No         & Yes         & No            & No         \\
    AWStats             & Partially  & Yes         & Partially\dag & Yes        \\
    Throughput monitor  & N/A        & No          & No            & No         \\
    Anteater            & No         & No          & No            & No         \\
    YAALA               & No         & Yes\ddag    & Yes           & No         \\
    Logparser/Lire      & Yes        & Yes         & Yes           & Yes        \\
    Logrep              & No         & Yes         & No            & No         \\
    Log Mail Analyser   & No         & Yes         & Partially     & No         \\ 
    \parsername{}       & Yes        & Yes\dag\dag & No            & Yes        \\
\end{tabular}

\dag AWStats can be configured to omit a section from a report, but not to
add extra sections.

@@

While there are other programs available which perform basic Postfix log
parsing (some to a greater level of detail than others), few attempt to
correlate log lines by queueid (none correlate by \pid{}) to produce an
overall record of the journey each mail traverses through Postfix.  None of
the reviewed parsers collect the breadth of information gathered by
\parsername{}, nor make it as easy to extend the parser to handle new log
lines.  Most other parsers immediately generate a report and discard the
data extracted from the logs; those which don't discard the data retain it
in a format inaccessible to other tools.  All of the parsers reviewed can
produce a report of greater or lesser detail and complexity, a facility not
offered by \parsername{}; reporting is deferred to a subsequent
application.  Table \refwithpage{Summary of parsers' features} provides a
summary of the parsers' features.

>>>>

Splunk aims to index all an organisation's logs, providing a centralised
view capable of searching and correlating diverse log sources.  The web
interface supports complicated searches, providing statistics and graphs in
real time, a facility not provided by \parsername{} (report generation has
been deferred to a subsequent program).  Saved searches\footnote{The author
was unable to save searches, though that may have been due to limitations
in the free version.} can be run periodically and the results emailed to a
recipient or sent to a shell script, which presumably can publish the
results as required (though possibly without the graphs and detailed
statistics); \parsername{} provides the database and leaves it to the user
to utilise it in any way, whenever they want, with no restrictions on usage
of the data.  The interface is optimised for interactive rather than
automated queries and it does not appear to be possible to write
independent tools to utilise the Splunk database; there are additional
reports available on \url{http://www.splunkbase.com/}, but the author was
unable to find any: every category was empty, even those which the
interface claimed had reports available.  Many types of reports are bundled
with the software, though most are variations of a bar or pie chart, with
the exception of bubble and heatmap graphs.  It is very easy to drill down
through the graphs to extract a portion of the data (e.g.\ select the hour
with the largest number of events, then select a particular host, and
finally a specific address), though it is not possible to search on partial
words.

@@

Copious documentation is made available on \url{http://www.splunk.com/},
but poor organisation and sheer abundance makes it extremely hard to find
useful information.  Searching using the website's interface confusingly
tends to return results from old documentation rather than new.  In
general the documentation appears to have been written by someone
intimately acquainted with the software and having difficulty understanding
how a newcomer would approach tasks or the questions they would ask.

@@

\subsubsection{Throughput monitor}

@@

\url{http://home.uninet.ee/~ragnar/throughput_monitor/} \newline (Last
checked 2008/05/01.)

@@

YAALA uses a plugin based system to analyse log files and produce \HTML{}
output reports, with all the parsing and report generation handled by
modules.  Using YAALA as a base would be only slightly less work than
starting from scratch, as both input and output modules would need to be
written specially; it may even be more work to implement the parser within
the constraints of YAALA\@.  YAALA supports storing previously gathered
data using Perl's Storable module~\cite{perl-storable}, so with enough
knowledge of the data structure YAALA stores it should be possible to
extract data with another Perl program also using the Storable module;
\parsername{} uses a well documented database which is accessible from the
majority of programming languages.  This information was gleaned from the
source code, as the documentation is sadly lacking.

@@

\subsubsection{Lire}

@@

\begin{table}[ht]
\begin{tabular}{l|l|l|l|l|l}
    Parser         & Parsed     & Data        & Custom        & Documentation? & Source \\
                   & test logs? & store?      & reports?      &                & code?  \\
    \hline         &            &             &               &                &        \\
    Pflogsumm      & Yes        & No          & Partial\dag   & Good           & Yes    \\
    Sawmill        & Yes        & Yes         & Searches      & Reasonable     & No     \\
    Splunk         & Yes        & Yes         & Searches      & Abundant       & No     \\
                   &            &             & \& reports    & but poor       &        \\
    Isoqlog        & No         & Yes         & No            & No             & Yes    \\
    AWStats        & Partially  & Yes         & Partial\dag   & Good           & Yes    \\
    Throughput     & N/A        & No          & No            & Brief but      & Yes    \\
    \tab{}monitor  &            &             &               & good           &        \\
    Anteater       & No         & No          & No            & No             & Yes    \\
    YAALA          & No         & Yes\ddag    & Searches      & Poor           & Yes    \\
    Lire           & Yes        & Yes         & Yes           & Yes            & Yes    \\
    Logrep         & No         & Yes         & No            & Barely         & Yes    \\
    \LMA{}         & No         & Yes         & No            & No             & Yes    \\ 
    \parsername{}  & Yes        & Yes\dag\dag & No            & Yes            & Yes    \\
\end{tabular}

\dag Sections can be omitted from a report, but extra sections can not be
added.

@@

While there are other programs available which perform basic Postfix log
parsing (some to a greater level of detail than others), few attempt to
correlate log lines by queueid (none correlate by \pid{}) to produce an
overall record of the journey each mail traverses through Postfix.  None of
the reviewed parsers collect the breadth of information gathered by
\parsername{}, nor make it as easy to extend the parser to handle new log
lines.  Most other parsers immediately generate a report and discard the
data extracted from the logs; those which don't discard the data retain it
in a format inaccessible to other tools.  Nearly all of the parsers
reviewed can produce a report of greater or lesser detail and complexity, a
facility not offered by \parsername{}; reporting is deferred to a
subsequent application.  The quality of the documentation offered by the
subset of parsers which provide some varies from unusable to good; none of
the parsers provide any documentation on the format of their data stores
(if any).  Less than half of the parsers were capable of parsing the
\numberOFlogFILES{} test log files, and improving or extending the parsers
would have been quite a difficult task for any of the parsers.  Table
\refwithpage{Summary of parsers' features} provides a summary of the
parsers' features.

### 2008-05-01 doc/logparser-parsing-rules.tex
Added cluster_groups to the rules table description and example rule. Improved Splunk review. Improved summary table in review of other parsers conclusion. Minor improvements in parser reviews. Improved review of other parsers conclusion.

\textbf{Field}      & \textbf{Value}                                    \\
name                & Unknown sender domain                             \\
description         & We do not accept mail from unknown domains        \\
restriction\_name   & reject\_unknown\_sender\_domain                   \\
postfix\_action     & REJECTED                                          \\
program             & \daemon{smtpd}                                    \\
regex               & \verb!^__RESTRICTION_START__ <(__SENDER__)>: !    \\
                    & \verb!Sender address rejected: Domain not found;! \\
                    & \verb!from=<\5> to=<(__RECIPIENT__)> !            \\
                    & \verb!proto=E?SMTP helo=<(__HELO__)>$!            \\
result\_cols        & recipient = 6; sender = 5                         \\
connection\_cols    & helo = 7                                          \\
result\_data        &                                                   \\
connection\_data    &                                                   \\
action              & REJECTION                                         \\
queueid             & 1                                                 \\
hits                & 0                                                 \\
hits\_total         & 0                                                 \\
priority            & 0                                                 \\

@@

>>>>

\textbf{Field}      & \textbf{Value}                                    \\
name                & Unknown sender domain                             \\
description         & We do not accept mail from unknown domains        \\
restriction\_name   & reject\_unknown\_sender\_domain                   \\
postfix\_action     & REJECTED                                          \\
program             & \daemon{smtpd}                                    \\
regex               & \verb!^__RESTRICTION_START__ <(__SENDER__)>: !    \\
                    & \verb!Sender address rejected: Domain not found;! \\
                    & \verb!from=<\5> to=<(__RECIPIENT__)> !            \\
                    & \verb!proto=E?SMTP helo=<(__HELO__)>$!            \\
result\_cols        & recipient = 6; sender = 5                         \\
connection\_cols    & helo = 7                                          \\
result\_data        &                                                   \\
connection\_data    &                                                   \\
action              & REJECTION                                         \\
queueid             & 1                                                 \\
hits                & 0                                                 \\
hits\_total         & 0                                                 \\
priority            & 0                                                 \\
cluster\_group      & 400                                               \\

@@

    \item [cluster\_group] The cluster\_group attribute is used by the
        Decision Tree algorithm described in a separate document; the
        parser does not use it in any way.
